---
title: "IDS investigation worksheet"
author: "by YDTeam: Cecilia, Frank, Junkai, Zhandos"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)

adjudications <- read_csv("data/CSV/AdjudicationsQ12023.csv")




```


```{r load-data}
library(tidyverse)
library(janitor)
 
adjudications <- adjudications %>%
  clean_names()

adjudications_clean <- adjudications %>%
  drop_na()


adjudications_clean <- adjudications_clean %>%
  rename(quarter = date)


glimpse(adjudications_clean)
colSums(is.na(adjudications_clean))


```

```{r}
library(dplyr)
library(ggplot2)


adjudications_clean %>%
  filter(sex == "M") %>% 
  group_by(offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 6) %>%
  ggplot(aes(x = reorder(offence, total), y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 6 Offences for Males (Q1 2023)",
    x = "Offence",
    y = "Total cases"
  )

adjudications_clean %>%
  filter(sex == "F") %>% 
  group_by(offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 6) %>%
  ggplot(aes(x = reorder(offence, total), y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 6 Offences for Females (Q1 2023)",
    x = "Offence",
    y = "Total cases"
  )

```

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(forcats)
library(scales)
library(gt)
library(glue)
library(janitor)

adj <- read_csv("AdjudicationsQ12023.csv") |> clean_names()

adj_clean <- adj |>
  mutate(
    count = suppressWarnings(as.numeric(count)),
    age_group = as.character(age_group),
    offence   = as.character(offence)
  ) |>
  filter(!is.na(age_group), !is.na(offence), !is.na(count), count > 0)

age_order <- c("15 - 17","18 - 20","21 - 24","25 - 29","30 - 39",
               "40 - 49","50 - 59","60 - 69","70 and over")
adj_clean <- adj_clean |>
  mutate(age_group = fct_relevel(age_group, age_order, after = 0))

age_offence_counts <- adj_clean |>
  group_by(age_group, offence) |>
  summarise(total = sum(count), .groups = "drop")

age_totals <- age_offence_counts |>
  group_by(age_group) |>
  summarise(age_total = sum(total), .groups = "drop")

age_offence_share <- age_offence_counts |>
  left_join(age_totals, by = "age_group") |>
  mutate(pct = total / age_total)

p_counts <- age_offence_counts |>
  ggplot(aes(x = total, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    data = age_totals,
    aes(x = age_total, y = age_group, label = comma(age_total)),
    nudge_x = max(age_totals$age_total, na.rm = TRUE) * 0.02, 
    inherit.aes = FALSE, size = 3
  ) +
  labs(
    title = "Adjudications by Age Group and Offence (Q1 2023)",
    x = "Total Cases",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)

p_share_labels <- age_offence_share |>
  mutate(lbl = ifelse(pct >= 0.05, percent(pct, accuracy = 1), "")) |>
  ggplot(aes(x = pct, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    aes(label = lbl),
    position = position_stack(vjust = 0.5),
    size = 3, color = "white"
  ) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title = "Composition of Offence Types within Each Age Group (Q1 2023)",
    x = "Share within Age Group",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)
```