---
title: "IDS investigation worksheet"
author: "by YDTeam: Cecilia, Frank, Junkai, Zhandos"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)

adjudications <- read_csv("data/CSV/AdjudicationsQ12023.csv")




```


```{r load-data}
library(tidyverse)
library(janitor)
 
adjudications <- adjudications %>%
  clean_names()

adjudications_clean <- adjudications %>%
  drop_na()


adjudications_clean <- adjudications_clean %>%
  rename(quarter = date)


glimpse(adjudications_clean)
colSums(is.na(adjudications_clean))


```
Topic 1：
Crime Patterns Across Age Groups in UK Prisons (Q1 2023)
Question:
What kinds of crimes are different age groups committing?
introduction:
Understanding how offence patterns vary across age groups is important for identifying behavioural trends within prisons.
In the Adjudications dataset (Q1 2023), prisoners are classified into age groups, and each adjudication is associated with an offence category such as Disobedience/disrespect, Unauthorized transactions, or Violence.
This section examines which age groups are responsible for particular types of offences and whether certain offences are more common among specific age ranges.
```{r}
library(dplyr)
library(ggplot2)


adjudications_clean %>%
  group_by(offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 6) %>%
  ggplot(aes(x = reorder(offence, total), y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 6 offences (Q1 2023)",
    x = "Offence",
    y = "Total cases"
  )


```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(forcats)
library(scales)
library(gt)
library(glue)
library(janitor)


adj <- read_csv("AdjudicationsQ12023.csv") |> clean_names()


adj_clean <- adj |>
  rename(
    age_group = age_group,
    offence   = offence,
    count     = count
  ) |>
  mutate(
    count = suppressWarnings(as.numeric(count)),
    age_group = as.character(age_group),
    offence   = as.character(offence)
  ) |>
  filter(!is.na(age_group), !is.na(offence), !is.na(count), count > 0)


age_order <- c("15 - 17","18 - 20","21 - 24","25 - 29","30 - 39","40 - 49","50 - 59","60 - 69","70 and over")
adj_clean <- adj_clean |>
  mutate(age_group = fct_relevel(age_group, age_order, after = 0))

age_offence_counts <- adj_clean |>
  group_by(age_group, offence) |>
  summarise(total = sum(count), .groups = "drop")


age_totals <- age_offence_counts |>
  group_by(age_group) |>
  summarise(age_total = sum(total), .groups = "drop")


age_offence_share <- age_offence_counts |>
  left_join(age_totals, by = "age_group") |>
  mutate(pct = total / age_total)
top_offences <- adj_clean |>
  group_by(offence) |>
  summarise(total_cases = sum(count), .groups = "drop") |>
  arrange(desc(total_cases))

top_offences |>
  mutate(
    Rank  = row_number(),
    total_cases = comma(total_cases)
  ) |>
  select(Rank, offence, total_cases) |>
  gt() |>
  tab_header(title = md("**Top offences (Q1 2023)**")) |>
  cols_label(
    offence = "Offence",
    total_cases = "Total cases"
  )
top3_by_age <- age_offence_share |>
  arrange(age_group, desc(total)) |>
  group_by(age_group) |>
  slice_head(n = 3) |>
  mutate(share = percent(pct, accuracy = 0.1)) |>
  ungroup() |>
  select(age_group, offence, total, share)

top3_by_age |>
  mutate(total = comma(total)) |>
  gt(groupname_col = "age_group") |>
  tab_header(title = md("**Top-3 offence types by age group**")) |>
  cols_label(
    offence = "Offence",
    total   = "Cases",
    share   = "Share"
  )
p_counts <- age_offence_counts |>
  ggplot(aes(x = total, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    data = age_totals,
    aes(x = age_total, y = age_group, label = comma(age_total)),
    nudge_x = max(age_totals$age_total, na.rm = TRUE) * 0.02, 
    inherit.aes = FALSE, size = 3
  ) +
  labs(
    title = "Adjudications by Age Group and Offence (Q1 2023)",
    x = "Total Cases",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)

p_counts
# ggsave("age_offence_counts.png", p_counts, width = 9, height = 6, dpi = 300)

age_offence_share <- age_offence_counts |>
  dplyr::group_by(age_group) |>
  dplyr::mutate(age_total = sum(total),
                pct = total / age_total) |>
  dplyr::ungroup()


p_share_labels <- age_offence_share |>
  dplyr::mutate(lbl = ifelse(pct >= 0.05, scales::percent(pct, accuracy = 1), "")) |>
  ggplot(aes(x = pct, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    aes(label = lbl),
    position = position_stack(vjust = 0.5),
    size = 3, color = "white"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Composition of Offence Types within Each Age Group (Q1 2023)",
    x = "Share within Age Group",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)

p_share_labels
# ggsave("age_offence_share.png", p_share_labels, width = 9, height = 6, dpi = 300)

```
```{r}

```


```{r}
adjudications_clean %>%
  group_by(sex) %>%
  summarise(total = sum(count)) %>%
  ggplot(aes(x = sex, y = total, fill = sex)) +
  geom_col() +
  labs(
    title = "Adjudications by Sex (Q1 2023)",
    x = "Sex",
    y = "Total Cases"
  )
```

```{r}

library(dplyr)
library(ggplot2)
library(stringr)

load("./data/CSV/punishments.rdata")

pun_clean <- punishments %>%
  mutate(
    no_days = as.numeric(`No of days`),
    age_mid = sapply(str_extract_all(`Age group`, "\\d+"),
                     function(v) if (length(v) >= 2) mean(as.numeric(v))
                                 else as.numeric(v[1]))
  ) %>%
  filter(!is.na(no_days), no_days > 0) %>%
  select(Sex, age_mid, no_days)

ggplot(pun_clean, aes(x = age_mid, y = no_days, color = Sex)) +
  geom_point(alpha = 0.25, size = 1.5) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, size = 1.2) +
  labs(
    title = "Punishment Length vs Age (Q1 2023)",
    x = "Age (years)",
    y = "Punishment Length (days)",
    color = "Sex"
  ) +
  theme_minimal(base_size = 14)


```

```{r}
#按监狱统计处罚数量
adjudications_clean %>%
  group_by(establishment) %>%
  summarise(total = sum(count)) %>%
  slice_max(total, n = 10) %>%
  ggplot(aes(x = reorder(establishment, total), y = total)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Prisons by Adjudications (Q1 2023)",
       x = "Prison",
       y = "Total disciplinary cases")

```


```{r}

```