---
title: "IDS investigation worksheet"
author: "by YDTeam: Cecilia, Frank, Junkai, Zhandos"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)

adjudications <- read_csv("data/CSV/AdjudicationsQ12023.csv")




```


```{r load-data}
library(tidyverse)
library(janitor)
 
adjudications <- adjudications %>%
  clean_names()

adjudications_clean <- adjudications %>%
  drop_na()


adjudications_clean <- adjudications_clean %>%
  rename(quarter = date)


glimpse(adjudications_clean)
colSums(is.na(adjudications_clean))


```
Topic 1：
Crime Patterns Across Age Groups in UK Prisons (Q1 2023)
Question:
What kinds of crimes are different age groups committing?
introduction:
Understanding how offence patterns vary across age groups is important for identifying behavioural trends within prisons.
In the Adjudications dataset (Q1 2023), prisoners are classified into age groups, and each adjudication is associated with an offence category such as Disobedience/disrespect, Unauthorized transactions, or Violence.
This section examines which age groups are responsible for particular types of offences and whether certain offences are more common among specific age ranges.
```{r}
library(dplyr)
library(ggplot2)


adjudications_clean %>%
  group_by(offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 6) %>%
  ggplot(aes(x = reorder(offence, total), y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 6 offences (Q1 2023)",
    x = "Offence",
    y = "Total cases"
  )


```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(forcats)
library(scales)
library(gt)
library(glue)
library(janitor)


adj <- read_csv("AdjudicationsQ12023.csv") |> clean_names()


adj_clean <- adj |>
  rename(
    age_group = age_group,
    offence   = offence,
    count     = count
  ) |>
  mutate(
    count = suppressWarnings(as.numeric(count)),
    age_group = as.character(age_group),
    offence   = as.character(offence)
  ) |>
  filter(!is.na(age_group), !is.na(offence), !is.na(count), count > 0)


age_order <- c("15 - 17","18 - 20","21 - 24","25 - 29","30 - 39","40 - 49","50 - 59","60 - 69","70 and over")
adj_clean <- adj_clean |>
  mutate(age_group = fct_relevel(age_group, age_order, after = 0))

age_offence_counts <- adj_clean |>
  group_by(age_group, offence) |>
  summarise(total = sum(count), .groups = "drop")


age_totals <- age_offence_counts |>
  group_by(age_group) |>
  summarise(age_total = sum(total), .groups = "drop")


age_offence_share <- age_offence_counts |>
  left_join(age_totals, by = "age_group") |>
  mutate(pct = total / age_total)
top_offences <- adj_clean |>
  group_by(offence) |>
  summarise(total_cases = sum(count), .groups = "drop") |>
  arrange(desc(total_cases))

top_offences |>
  mutate(
    Rank  = row_number(),
    total_cases = comma(total_cases)
  ) |>
  select(Rank, offence, total_cases) |>
  gt() |>
  tab_header(title = md("**Top offences (Q1 2023)**")) |>
  cols_label(
    offence = "Offence",
    total_cases = "Total cases"
  )
top3_by_age <- age_offence_share |>
  arrange(age_group, desc(total)) |>
  group_by(age_group) |>
  slice_head(n = 3) |>
  mutate(share = percent(pct, accuracy = 0.1)) |>
  ungroup() |>
  select(age_group, offence, total, share)

top3_by_age |>
  mutate(total = comma(total)) |>
  gt(groupname_col = "age_group") |>
  tab_header(title = md("**Top-3 offence types by age group**")) |>
  cols_label(
    offence = "Offence",
    total   = "Cases",
    share   = "Share"
  )
p_counts <- age_offence_counts |>
  ggplot(aes(x = total, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    data = age_totals,
    aes(x = age_total, y = age_group, label = comma(age_total)),
    nudge_x = max(age_totals$age_total, na.rm = TRUE) * 0.02, 
    inherit.aes = FALSE, size = 3
  ) +
  labs(
    title = "Adjudications by Age Group and Offence (Q1 2023)",
    x = "Total Cases",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)

p_counts
# ggsave("age_offence_counts.png", p_counts, width = 9, height = 6, dpi = 300)

age_offence_share <- age_offence_counts |>
  dplyr::group_by(age_group) |>
  dplyr::mutate(age_total = sum(total),
                pct = total / age_total) |>
  dplyr::ungroup()


p_share_labels <- age_offence_share |>
  dplyr::mutate(lbl = ifelse(pct >= 0.05, scales::percent(pct, accuracy = 1), "")) |>
  ggplot(aes(x = pct, y = age_group, fill = offence)) +
  geom_col() +
  geom_text(
    aes(label = lbl),
    position = position_stack(vjust = 0.5),
    size = 3, color = "white"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Composition of Offence Types within Each Age Group (Q1 2023)",
    x = "Share within Age Group",
    y = "Age Group",
    fill = "Offence"
  ) +
  theme_minimal(base_size = 12)

p_share_labels
# ggsave("age_offence_share.png", p_share_labels, width = 9, height = 6, dpi = 300)

```

Topic 2：
Punishment Length vs Age and Sex
Question:
How does punishment length vary between males and females across different age groups?
introduction:
In this section, we examine whether males and females receive punishments of different lengths, and whether age plays a role in these differences.
We use the PunishmentsQ12023 dataset to visualize the relationship between punishment length and age for both sexes.
```{r}
adjudications_clean %>%
  group_by(sex) %>%
  summarise(total = sum(count)) %>%
  ggplot(aes(x = sex, y = total, fill = sex)) +
  geom_col() +
  labs(
    title = "Adjudications by Sex (Q1 2023)",
    x = "Sex",
    y = "Total Cases"
  )
```

```{r}
library(dplyr)
library(ggplot2)
library(broom)

#pun_clean <- read.csv("C:/Users/cnmon/Documents/YDteam/data/CSV/PunishmentsQ12023.csv", header = TRUE)
 
load("C:/Users/cnmon/Documents/YDteam/data/CSV/punishments.rdata")
pun_clean <- punishments
pun_clean <- pun_clean %>%
  mutate(
    Sex = factor(Sex),                
    no_days = as.numeric(`No of days`),    
    age_mid = as.character(`Age group`)
  ) %>%
  filter(!is.na(no_days), !is.na(age_mid))


mod_main <- lm(no_days ~ Sex + age_mid, data = pun_clean)

cat("\n=== Main model: no_days ~ Sex + age_mid ===\n")
summary(mod_main)
tidy(mod_main)


p_main <- ggplot(pun_clean, aes(x = age_mid, y = no_days, color = Sex)) +
  geom_point(alpha = 0.35, size = 1.6) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = 1.1) +
  labs(
    title = "Punishment Length vs Age (Main Effects)",
    subtitle = "Separate OLS fits by Sex (visualizing main effects)",
    x = "Age (midpoint, years)",
    y = "Punishment length (days)",
    color = "Sex"
  ) +
  theme_minimal(base_size = 13)
p_main

mod_inter <- lm(no_days ~ Sex * age_mid, data = pun_clean)

cat("\n=== Interaction model: no_days ~ Sex * age_mid ===\n")
summary(mod_inter)
tidy(mod_inter)


p_inter <- ggplot(pun_clean, aes(x = age_mid, y = no_days, color = Sex)) +
  geom_point(alpha = 0.35, size = 1.6) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = 1.1) +
  labs(
    title = "Punishment Length vs Age (Interaction Model)",
    subtitle = "Lines are OLS fits by Sex; slope difference = interaction",
    x = "Age (midpoint, years)",
    y = "Punishment length (days)",
    color = "Sex"
  ) +
  theme_minimal(base_size = 13)
p_inter



p_jitter_ci <- ggplot(pun_clean, aes(x = age_mid, y = no_days, color = Sex, fill = Sex)) +
  geom_jitter(width = 0.8, height = 0.2, alpha = 0.25, size = 1.6, shape = 16) +  
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 1.2, alpha = 0.2) +
  coord_cartesian(xlim = c(15, 60), ylim = c(0, 20)) +                              
  labs(
    title    = "Punishment Length vs Age (by Sex, with 95% CI)",
    subtitle = "Shaded areas show 95% confidence intervals; most punishments <20 days",
    x = "Age (midpoint of age group, years)",
    y = "Punishment Length (days)",
    color = "Sex",
    fill  = "Sex"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")

p_jitter_ci
# ggsave("punishment_jitter_CI.png", p_jitter_ci, width = 9, height = 6, dpi = 300)

```

```{r}
#按监狱统计处罚数量
adjudications_clean %>%
  group_by(establishment) %>%
  summarise(total = sum(count)) %>%
  slice_max(total, n = 10) %>%
  ggplot(aes(x = reorder(establishment, total), y = total)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Prisons by Adjudications (Q1 2023)",
       x = "Prison",
       y = "Total disciplinary cases")

```


```{r}
#new topic
```

```{r}

```
