---
title: "IDS investigation worksheet"
author: "by YDTeam: Cecilia, Frank, Junkai, Zhandos"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)

adjudications <- read_csv("data/CSV/AdjudicationsQ12023.csv")




```


```{r load-data}
library(tidyverse)
library(janitor)
 
adjudications <- adjudications %>%
  clean_names()

adjudications_clean <- adjudications %>%
  drop_na()


adjudications_clean <- adjudications_clean %>%
  rename(quarter = date)


glimpse(adjudications_clean)
colSums(is.na(adjudications_clean))


```

```{r}
#按监狱统计处罚数量
adjudications_clean %>%
  group_by(establishment) %>%
  summarise(total = sum(count)) %>%
  slice_max(total, n = 10) %>%
  ggplot(aes(x = reorder(establishment, total), y = total)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Prisons by Adjudications (Q1 2023)",
       x = "Prison",
       y = "Total disciplinary cases")


```

```{r}

```

```{r}
adjudications_clean %>%
  group_by(sex) %>%
  summarise(total = sum(count)) %>%
  ggplot(aes(x = sex, y = total, fill = sex)) +
  geom_col() +
  labs(
    title = "Adjudications by Sex (Q1 2023)",
    x = "Sex",
    y = "Total Cases"
  )

```

```{r}

library(dplyr)
library(ggplot2)


adjudications_clean %>%
  group_by(offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 6) %>%
  ggplot(aes(x = reorder(offence, total), y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 6 offences (Q1 2023)",
    x = "Offence",
    y = "Total cases"
  )


```

```{r}
library(dplyr)
library(ggplot2)


age_offence_counts <- adjudications_clean %>%
  group_by(age_group, offence) %>%
  summarise(total = sum(count, na.rm = TRUE), .groups = "drop") %>%
  group_by(age_group) %>%
  mutate(age_total = sum(total)) %>%
  ungroup()

age_offence_counts %>%
  ggplot(aes(x = age_group, y = total, fill = offence)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Adjudications by Age Group and Offence (Q1 2023)",
    x = "Age Group",
    y = "Total Cases"
  )
```


```{r}

# separate males and females
adjudications_clean %>% 
  group_by(gender) %>% 

data_distance <- filter(data_distance, !is.na(runner_type))

data_distance <- ungroup(data_distance) %>% 
  mutate(
    distance_km = distance..m./1000,
    avg_speed = distance..m./elapsed.time..s.
  )
  
data_distance %>% 
  ggplot(
    aes(x = distance_km, y = avg_speed, color = runner_type)
  ) +
  geom_point(alpha = 0.25) +
  #adding smooth lines
  geom_smooth(method = lm, se = FALSE) +
  scale_color_manual(values = c("long" = "orange", "short" = "forestgreen"), labels = c("long" = "Long-Distance Runner", "short" = "Short-Distance Runner")) +
  theme(legend.title = element_blank()) +
  labs(
    x = "Run distance (km)",
    y = "Run average speed (m/s)",
    title = "Speed vs Distance for Long- and Short-Distance Runners",
    subtitle = "Grouped by runners' average distance: >13km vs <7km"
  )
    
```{r}
library(readxl)
library(dplyr)
library(ggplot2)


load("data/CSV/punishments.rdata")

colnames(punishments)

pun_clean <- punishments %>%
  mutate(
    no_days = as.numeric(`No of days`)   
  ) %>%
  filter(
    !is.na(no_days),                  
    no_days > 0                        
  ) %>%
  select(Sex, `Age group`, no_days)
ggplot(pun_clean, aes(x = `Age group`, y = no_days, color = Sex)) +
  geom_jitter(alpha = 0.4, width = 0.2) +  
  geom_point(alpha=0.02) +
  geom_smooth(method = 'lm', se = FALSE, size = 1.2) +  
  labs(
    title = "Punishment Lengths by Sex Across Age Groups (Q1 2023)",
    x = "Age Group",
    y = "Punishment Length (days)",
    color = "Sex"
  ) +
  theme_minimal(base_size = 14)



```